---
title: "Why has number of car casualties in Slovenia fallen in last years?"
author: "Elvira Nassirova"
date: "28 Jan 2018"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)

library(dplyr)
library(ggplot2)
library(plotly)
library(readxl)
library(tidyr)
library(RMySQL)
library(knitr)
library(kableExtra)
```


```{r slvdata}

acc_data <- read.csv('data/Accidents_data.txt', sep = '\t', stringsAsFactors = F)

slv_data <- filter(acc_data, Country == 'Slovenia' & measure == 'NBR') %>% 
                  arrange(subject, measure) %>% 
            mutate(subject = ordered(subject, 
                                     levels = c('INJURE', 'ACCIDENTCASUAL', 'DEATH')))

#Plot 1. Road accidents Statstics in Slovenia, 2000 - 2006=====
p1 <- ggplot(slv_data, aes(x = time, y = value, color = subject, group = subject)) +
  geom_point(size=3)+
  geom_line(size=1)+
  #geom_text(aes(label=value), size=3, vjust = -1.5, 
  #          col='black', fontface="bold", show.legend = FALSE)+
  geom_vline(xintercept = 2008, size = 1, alpha = 0.6)+
  scale_x_continuous(name='', limits = c(2000, 2016), breaks = seq(2000, 2016, 2))+
  scale_y_continuous(name = 'Thousands of accidents')+
  scale_color_manual(name = 'Perspectives', 
                     labels = c('Injures', 'Accidents involving casualties', 'Deaths'),
                     values = c("#5586a6", "#fa990e", "#80a300"))+
  theme_bw()+
  theme(legend.position = 'bottom')+
  annotate("text", x = 2008, y = 5000, size = 3, 
           label = "Starting from 2008
numbers are falling")+
  labs(title = 'Road accidents Statstics in Slovenia, 2000 - 2006',
       caption = 'Source: OECD Data 
                  URL: https://data.oecd.org/transport/road-accidents.htm')

#Plot 2. Police Report Data =====
report_table <- read_xlsx('data/regulations_stat.xlsx') %>% 
                  mutate_at(c('2006', '2007', '2008', '2009'),
                            funs(format(., big.mark=" ")))

agg_acc_data <- acc_data %>% filter(measure == 'NBR') %>% 
                        group_by(Country, time) %>% summarise(accidents = sum(value))

immigration <- read_xlsx('data/Imm Directions SLOVENIA STAT.xlsx', sheet = 2) %>% 
                gather(key = 'time', value = 'immigrants', -Country) %>% 
                mutate(Country = ordered(Country, 
                                 levels = c('Bosnia-Herzegovina', 'Serbia', 'Macedonia')),
                       time = as.integer(time))
                #merge(agg_acc_data, by = c('Country', 'time'))

#Plot 3. Dynamics of immigration to Slovenia, 2008 - 2015 =====
p2 <- ggplot(immigration, aes(x = time, y = immigrants, color = Country, group = Country)) +
  geom_point(size=3)+
  geom_line(size=1)+
  scale_x_continuous(name='')+
  scale_y_continuous(name = 'Number of Immigrants')+
  scale_color_manual(values = c("#5586a6", "#fa990e", "#80a300"))+
  theme_bw()+
  theme(legend.position = 'bottom')+
  labs(title = 'Dynamics of immigration to Slovenia, 2008 - 2015',
       caption = 'Source: Statistical Office of the Republic of Slovenia')

total_imm <- immigration %>% group_by(time) %>% summarise(immigrants = sum(immigrants)) %>% 
                merge(filter(agg_acc_data, Country == 'Slovenia'), by = 'time')

#Plot 4. Accidents and Immigration correlation =====
p3 <- ggplot(total_imm, aes(x = accidents, y = immigrants))+
  geom_point(shape = 21, colour = "black", fill = "#5586a6", size = 3, stroke = 1.5)+
  geom_smooth(method='lm')+
  geom_text(aes(label=time), size=3, vjust = -1.5, 
            col='black', fontface="bold", show.legend = FALSE)+
  ylab('Number of Immigrants')+
  scale_x_continuous(name = 'Number of accidents in Slovenia', 
                     limits = c(14000, 22000), breaks = seq(14000, 22000, 2000))+
  annotate("text", x = 18000, y = 20000, size = 5, 
           label = paste('R^2 == ', 
                          round(cor(total_imm$immigrants, total_imm$accidents)^2, 2)),
           parse=TRUE)+
  labs(title = 'Accidents and Immigration correlation')
```

As we can see from the graph below there is a significant fall in a number of car accidents starting from 2008.

```{r}
p1
```

The first thing we are going to consult is [Slovenian Police reports](https://www.policija.si/eng/index.php/statistics). According to them, in 2008 there was an introduction of harsher penalties through the amended Road Safety Act. Indeed, a lot less violations were conducted in 2008, mainly due to the Road Traffic Safety Act.

```{r}
kable(report_table, caption = 'Violations detected during road traffic controls, 2006 - 2009', align = c('l', rep('r', 4)))
```

The second interesting thing is that starting from 2010 there is a big change in net migration in Slovenia mostly due to fall of immigrants from Bosnia-Herzegovina, Serbia and Macedonia. In addition, a decrease in a number of accidents from 2009 to 2010 was seen in Slovenia (see the first graph).

```{r}
p2
```

According to the [World Health Organization report](http://www.who.int/violence_injury_prevention/road_safety_status/2015/GSRRS2015_data/en/) estimated road traffic death rate per 100 000 population in these countries is higher (from 7.7 in Serbia to 17.7 in Bosnia-Herzegovina) than in Slovenia (6.4). If we build a scatter plot for a total number of immigrants from the indicated countries and a total number of car accidents in Slovenia, we can see correlation. 

```{r}
p3
```

Overall, we may conclude there are at least **two reasons**:

* New Road Safety Act
* Decreased number of immigrants
